The regularity of ODEs and thimble integrals \\ with respect to Borel summation

### 2024-04-06 exp-resurgence.tex
Revise intro to action on trans-series

%%It follows that $\borel_\zeta$ is a formal inverse of the $\laplace_{\zeta, 0}$: let $\varphi=\sum_{n=0}^{\infty}a_n \zeta^n$
%\begin{align*}
 %   \borel_\zeta \laplace_{\zeta}\varphi&=\borel_\zeta \left[\int_{\mathcal{J}_{\zeta,0}}e^{-z\zeta} \varphi d\zeta\right]\\
  %  &=\borel_\zeta\left[\sum_{n=0}^{\infty}a_n \int_{\mathcal{J}_{\zeta,0}} e^{-z\zeta}\zeta^n d\zeta\right]\\
  %  &=\borel_\zeta\left[\sum_{n=0}^{\infty}a_n z^{-n-1}n! \right]\\
  %  &=\sum_{n=0}^{\infty}a_n \frac{\zeta^n}{n!}
%\end{align*}
%where in the second step we formally interchange the integral and the sum. 
%\color{black}
\subsubsection{Action on trans-series}\label{sec:action_transseries}
%
\color{DarkTurquoise}
\color{RoyalBlue}
A trans-series is a generalization of usual power series, whose building blocks (called trans-monomials) are of the form $e^{\alpha_n z}\varphi_n$ where $\alpha_n$ is a sequence of complex numbers such that \textcolor{PaleVioletRed}{[exponents must lie along a ray]} $0=\alpha_0 e^{-i\theta}< \alpha_1 e^{-i\theta}<...<\alpha_n e^{-i\theta}<...$ and $\varphi_n\in\C\llbracket z^{-1}\rrbracket$ for every $n$:

@@

It's natural to ask whether we can extend the definition of $\borel_\zeta$ to transseries, and we will achieve it using the relationship between $\laplace_{\zeta, \alpha}$ and $\laplace_{\zeta_\alpha, 0}$ in identity~\eqref{change-chart}.
\color{black}
\begin{definition}
    The action of $\borel_\zeta$ on trans-monomials is the formal inverse of $\laplace_{\zeta,\alpha}$
    \[\laplace_{\zeta,\alpha}\borel_\zeta\big[e^{-z\alpha}z^{-n-1}\big]=e^{-z\alpha}z^{-n-1}\]
    %\begin{center}
%\begin{tikzcd}[every arrow/.append style={shift left}]
 %e^{-z\alpha}\{z^{-1}, z^{-2}, z^{-3}, z^{-4}, \ldots \} \arrow{d}{\borel_{\zeta}} \\ \{1, \zeta_\alpha, \zeta_\alpha^2, \zeta_\alpha^3, \ldots\} \arrow{u}{\laplace_{\zeta,\alpha}}
%\end{tikzcd}
%\end{center}
    and it extends by linearity to transseries. 
\end{definition}
In particular, from identity~\eqref{change-chart} we deduce 
\begin{align*}
    e^{-z\alpha}z^{-n-1}&=e^{-z\alpha}\laplace_{\zeta_\alpha,0}\borel_\zeta\big[e^{-\zeta\alpha}z^{-n-1}\big]\\
     z^{-n-1}&=\laplace_{\zeta_\alpha,0}\borel_\zeta\big[e^{-\zeta\alpha}z^{-n-1}\big]
\end{align*}
and taking the inverse of $\laplace_{\zeta_\alpha,0}$ we find
%Indeed it is possible by formalizing the change of variable identity that relates $\laplace_{\zeta, \alpha}$ to $\laplace_{\zeta_\alpha, 0}$ (see Section \ref{translation} identity \eqref{change-chart}).

@@

    \[\borel_{\zeta}\big[e^{-z\alpha}z^{-n-1}\big]=\frac{\zeta_\alpha^n}{n!}.\]

@@

Integrating both sides of equation~\eqref{eqn:diff-int-split}, we get
\[ \fracderiv{-1}{\zeta}{0}\;\tfrac{\partial}{\partial \zeta}\;f = -\left[ \fracderiv{-1}{\zeta}{0} \circ h_0 + \sum_{k = 1}^m \fracderiv{-k-1}{\zeta}{0} \circ h_{-k} \right] \phi. \]
Since $\fracderiv{-1}{\zeta}{0}\;\tfrac{\partial}{\partial \zeta}$ acts as the identity on functions that vanish at $\zeta = 0$, this simplifies to
\[ f = -\left[ \sum_{k = 0}^m \fracderiv{-k-1}{\zeta}{0} \circ h_{-k} \right] \phi, \]
which rearranges to the second equation in the statement.
\end{proof}
\newpage
\bibliographystyle{plain}
\bibliography{airy-resurgence}
\end{document}

>>>>

%%It follows that $\borel_\zeta$ is a formal inverse of the $\laplace_{\zeta, 0}$: let $\varphi=\sum_{n=0}^{\infty}a_n \zeta^n$
%\begin{align*}
 %   \borel_\zeta \laplace_{\zeta}\varphi&=\borel_\zeta \left[\int_{\mathcal{J}_{\zeta,0}}e^{-z\zeta} \varphi d\zeta\right]\\
  %  &=\borel_\zeta\left[\sum_{n=0}^{\infty}a_n \int_{\mathcal{J}_{\zeta,0}} e^{-z\zeta}\zeta^n d\zeta\right]\\
  %  &=\borel_\zeta\left[\sum_{n=0}^{\infty}a_n z^{-n-1}n! \right]\\
  %  &=\sum_{n=0}^{\infty}a_n \frac{\zeta^n}{n!}
%\end{align*}
%where in the second step we formally interchange the integral and the sum. 
%\color{black}
\subsubsection{Action on trans-series}\label{sec:action_transseries}
%
\color{DarkBlue}
{\em Trans-series} are formal objects that greatly generalize formal power series. Their full definition is quite involved~\cite{aniceto2019primer}, especially when complex exponents are allowed~\cite{van-der-hoeven2001complex}. We will only see examples of the simple form
\[ \sum_{\alpha \in A} e^{-\alpha z} z^{\sigma_\alpha} \series{\varphi}_\alpha, \]
where $A$ is a finite set of complex numbers, each $\sigma_\alpha$ is a real number, and each $\series{\varphi}_\alpha$ is a formal power series in $\C\llbracket z^{-1} \rrbracket$. Each term $e^{-\alpha z} z^{\sigma_\alpha} \series{\varphi}_\alpha$ is called a {\em trans-monomial}.

Trans-series of this form play a crucial role in the theory of ODEs with irregular singularities. In particular, they exhibit a formal analogue of the Stokes phenomenon: their asymptotics can change sharply from one direction to the next. Trans-monomials $e^{-\alpha z} z^{\sigma_\alpha} \series{\varphi}_\alpha$ can be ordered by $\Re(e^{i\theta} \alpha)$ for any direction $\theta$. When a trans-series converges, its sum becomes dominated by the lowest term in this ordering as $\Re(e^{-i\theta} z)$ grows.

We can extend the definition of the Borel transform to trans-series by recognizing that $\laplace_{\zeta, \alpha}$ sends ordinary monomials to trans-monomials, and then taking advantage of the relationship between $\laplace_{\zeta, \alpha}$ and $\laplace_{\zeta_\alpha, 0}$ given by identity~\eqref{change-chart}.
\par\color{RoyalBlue}
A trans-series is a generalization of usual power series, whose building blocks (called trans-monomials) are of the form $e^{\alpha_n z}\varphi_n$ where $\alpha_n$ is a sequence of complex numbers such that \textcolor{PaleVioletRed}{[exponents must lie along a ray]} $0=\alpha_0 e^{-i\theta}< \alpha_1 e^{-i\theta}<...<\alpha_n e^{-i\theta}<...$ and $\varphi_n\in\C\llbracket z^{-1}\rrbracket$ for every $n$:

@@

It's natural to ask whether we can extend the definition of $\borel_\zeta$ to transseries, and we will achieve it using the relationship between $\laplace_{\zeta, \alpha}$ and $\laplace_{\zeta_\alpha, 0}$ in identity~\eqref{change-chart}.
\begin{definition}
On each ordinary monomial within a trans-monomial, $\borel_\zeta$ acts as the formal inverse of $\laplace_{\zeta,\alpha}$:
\[\laplace_{\zeta,\alpha}\borel_\zeta\big[e^{-\alpha z} z^{-n-1} \big]=e^{-\alpha z} z^{-n-1}. \]
The action of $\borel_\zeta$ extends to general trans-monomials and trans-series by countable linearity.
From this definition, and identity~\eqref{change-chart}, we deduce that
e^{-z\alpha}\laplace_{\zeta_\alpha,0}\borel_\zeta\big[e^{-\alpha z} z^{-n-1}\big] & = e^{-z\alpha} z^{-n-1} \\
\laplace_{\zeta_\alpha,0}\borel_\zeta\big[e^{-\alpha z} z^{-n-1}\big] & = z^{-n-1} \\
\borel_{\zeta}\big[e^{-z\alpha}z^{-n-1}\big] & = \frac{\zeta_\alpha^n}{n!}.
%%%and taking the inverse of $\laplace_{\zeta_\alpha,0}$ we find
%Indeed it is possible by formalizing the change of variable identity that relates $\laplace_{\zeta, \alpha}$ to $\laplace_{\zeta_\alpha, 0}$ (see Section \ref{translation} identity \eqref{change-chart}). 

@@

%%%\[\borel_{\zeta}\big[e^{-z\alpha}z^{-n-1}\big]=\frac{\zeta_\alpha^n}{n!}.\]

@@

Integrating both sides of equation~\eqref{eqn:diff-int-split}, we get
\[ \fracderiv{-1}{\zeta}{0}\;\tfrac{\partial}{\partial \zeta}\;f = -\left[ \fracderiv{-1}{\zeta}{0} \circ h_0 + \sum_{k = 1}^m \fracderiv{-k-1}{\zeta}{0} \circ h_{-k} \right] \phi. \]
Since $\fracderiv{-1}{\zeta}{0}\;\tfrac{\partial}{\partial \zeta}$ acts as the identity on functions that vanish at $\zeta = 0$, this simplifies to
\[ f = -\left[ \sum_{k = 0}^m \fracderiv{-k-1}{\zeta}{0} \circ h_{-k} \right] \phi, \]
which rearranges to the second equation in the statement.
\end{proof}
\newpage
\bibliographystyle{utphys}
\bibliography{airy-resurgence}
\end{document}
